A molecular taxonomy of tumors independent of tissue-of-origin

### 2020-07-20 text/manuscript.Rmd

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rstan)
library(foreach)
library(sqldf)
library(tidyverse)
library(ggthemes)
library(data.table)
library(biomaRt)
library(dplyr)
library(GenomicRanges)
library(SummarizedExperiment)
library(reactome.db)
library(plyr)
library(cluster)
library(pheatmap)
library(clusterProfiler)
library(org.Hs.eg.db)
library(umap)
library(dbscan)
library(class)
library(TCGAbiolinks)
library(patchwork)
library(ggplotify)
library(ComplexHeatmap)
library(knitr)
library(kableExtra)
library(UpSetR)
library(cowplot)
```

@@

```{r child='../code/figure1.Rmd'}
```

In order to classify tumors using this dataset, we used multiple correspondence
analysis (MCA) [@JSSv025i01]. First we determined the number of dimensions
containing useful information (**figure 1b**). Then we performed a UMAP
analysis, both in order to summarize the MCA graphically, and as a preprocessing
step to boost the performance of density based clustering. The resulting map was
notable for its lobed structure, with several large, robust projections
reproducible regardless of random seed setting. A representative version of this
UMAP is shown in **figure 1c** as a 2D projection. Following this spatial
mapping we attempted to define groupings of similar tumors within the spatial
map using HDBSCAN, which performs hierarchical clustering and provides metrics
of cluster stability and probabilities of cluster membership for each node.
However, HDBSCAN is sensitive to several parameters; key for our analysis are
the minimum number of tumor samples in a cluster that capture the maximum number
of tumor samples, measured by probability of membership of $\geq 5\%$ in at
least one class. Thus, we created a score metric as the fraction of classified
tumors with max probability $< 5\%$ in one class and chose a cluster size of 95
to minimize the score function (**figure 1d**). HDBSCAN with these settings
resulted in 12 distinct high-density clusters which we then projected onto the
UMAP (**figure 1e**). This classified <mark>x out of 7,607 tumors</mark> but
still resulted in a significant fraction of unclassified tumor samples. Since we
ultimately wish to be able to classify any tumor using this scheme, we performed
k Nearest Neighbors (kNN) analysis, which computes a similarity metric to every
tumor in the set and then lets the $k$ most similar tumors "vote" as to the
identity of the query tumor sample based on their classification labels. We set
$k$ to be the square root of the number of tumor samples (87). Using this
method, we assigned cluster membership to the remaining tumors (**figure 1f**).

@@

```{r child='../code/figure4.Rmd'}
```

@@

```{r child='table2.Rmd'}
```

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rstan)
library(foreach)
library(sqldf)
library(tidyverse)
library(ggthemes)
library(data.table)
library(biomaRt)
library(dplyr)
library(GenomicRanges)
library(SummarizedExperiment)
library(reactome.db)
library(plyr)
library(cluster)
library(pheatmap)
library(clusterProfiler)
library(org.Hs.eg.db)
library(dbscan)
library(class)
library(TCGAbiolinks)
library(patchwork)
library(ggplotify)
library(ComplexHeatmap)
library(knitr)
library(kableExtra)
library(UpSetR)
library(cowplot)
library(doRNG)
library(uwot)
library(janitor)
library(FactoMineR)
```

@@

```{r child='../code/figure1.Rmd'}
```
In order to classify tumors using this dataset, we used multiple correspondence
analysis (MCA) [@JSSv025i01]. First we determined the number of dimensions
containing useful information (**figure 1b**). Then we performed a UMAP
analysis, both in order to summarize the MCA graphically, and as a preprocessing
step to boost the performance of density based clustering. The resulting map was
notable for its lobed structure, with several large, robust projections
reproducible regardless of random seed setting. A representative version of this
UMAP is shown in **figure 1c** as a 2D projection. Following this spatial
mapping we attempted to define groupings of similar tumors within the spatial
map using HDBSCAN, which performs hierarchical clustering and provides metrics
of cluster stability and probabilities of cluster membership for each node.
However, HDBSCAN is sensitive to several parameters; key for our analysis are
the minimum number of tumor samples in a cluster that capture the maximum number
of tumor samples, measured by probability of membership of $\geq 5\%$ in at
least one class. Thus, we created a score metric as the fraction of classified
tumors with max probability $< 5\%$ in one class and chose a cluster size of 95
to minimize the score function (**figure 1d**). HDBSCAN with these settings
resulted in 12 distinct high-density clusters which we then projected onto the
UMAP (**figure 1e**). This classified <mark>x out of 7,607 tumors</mark> but
still resulted in a significant fraction of unclassified tumor samples. Since we
ultimately wish to be able to classify any tumor using this scheme, we performed
k Nearest Neighbors (kNN) analysis, which computes a similarity metric to every
tumor in the set and then lets the $k$ most similar tumors "vote" as to the
identity of the query tumor sample based on their classification labels. We set
$k$ to be the square root of the number of tumor samples (87). Using this
method, we assigned cluster membership to the remaining tumors (**figure 1f**).

@@

```{r child='../code/figure4_function.Rmd'}
```

@@

```{r child='../code/table2.Rmd'}
```
